Expression Profiles
-------------------

To get a sense for how the genes were clustered, let's plot plot a subset of
the expression profiles and highlight the genes in each module.

**Note**: In order to improve the appearance of the expression profiles,
samples have been averaged wihtin each condition to produce a single x-tick per
condition. This is _not_ what is done during the actual clustering used to
generate the network - in that case each replicate is kept as an individual
variable in the gene expression vectors used for clustering. This mean,
however, that there may be some cases where the expression profile
visualizations give the appearance of grouping together genes which seem very
different. This is likely the result of outlier samples pulling the
within-condition averages up or down relative to the center of the clusters.

For reference, below is an example expression profile plot for where samples
have *not* been averaged by condition.

```{r example_expr_plot}
# example module expression
module_color <- module_colors[sample(1:length(module_colors), 1)]
module_expr <- exprs(network_counts$final)[module_colors == module_color,]

# wide to long
dat <- melt(module_expr)
colnames(dat) <- c('gene_id', 'sample_id', 'log_cpm')
dat$sample_id <- factor(dat$sample_id)
#dat <- cbind(dat, condition=rep(condition, each=nrow(module_expr)))

ggplot(data=dat,
    aes(x=sample_id, y=log_cpm, group=gene_id), color='#333') +
    geom_line(size=0.2) +
    ggtitle(sprintf("Gene Expression Profiles (%s)", module_color)) +
    xlab("Sample") +
    ylab("Expression level (log2-CPM)") +
    theme(axis.text=element_text(colour="#333333"), 
            axis.text.x=element_text(angle=90, hjust=1, size=16))
```

### Combine replicates

<!--
QUESTION: Can the variance between replicates for a given gene tell us anything
useful that with respect to network construction? For example, perhaps genes
with less variance within replicates are more tightly regulated. On the flip
side, genes with a larger variance may just be noise and not as interesting in
the context of the network.
-->

```{r combine_replicates}
combined_counts <- combine_replicates(exprs(network_counts$final), condition)
combined_counts_table <- cbind(gene_id=rownames(network_counts$final),
                               combined_counts)

# convert to long
counts_long <- melt_counts(combined_counts, CONFIG$condition_mapping)
colnames(counts_long) <- c('gene_id', 'condition', 'log_cpm')

# add clustering results and convert to a factor
counts_long <- cbind(counts_long, cluster=as.factor(module_colors))

# save result
COEXPRESSION_NETWORK_RESULT$counts_long <- counts_long
```

```{r expression_profiles, fig.width=18, eval=TRUE, fig.show='hold', results='hide', message=FALSE}
# Next, highlight a single cluster at a time
# rather than displaying all of the clusters, let's just choose a few at
# random to give an idea of what the picture looks like.

# sort by order on dendrogram (left to right)
module_order <- unique(module_colors[gene_tree$order])

# TODO: deal with warnings

## 1) Warning in if (is.na(module_order)) {: the condition has length > 1 and
## only the first element will be used

# 2) Aggregation function missing: defaulting to length (?)

if (CONFIG$include_plots) {
    module_expression_profile_plot(counts_long, module_colors, module_order,
                                   max_plots=3)
}
```

