KEGG Enrichment analysis
------------------------

Other tools to consider in the future:

- [NetPathMiner](http://bioconductor.org/packages/release/bioc/html/NetPathMiner.html)
- [ReactomePA](http://bioconductor.org/packages/release/bioc/html/ReactomePA.html)
- [KEGGprofile](http://www.bioconductor.org/packages/release/bioc/html/KEGGprofile.html)
- [pathview](http://www.bioconductor.org/packages/release/bioc/html/pathview.html)

First, create a mapping from gene ids to KEGG pathways, similar to what we did
above with GO terms.

Note that in earlier versions of this analyses, some KEGG identifiers not
corresponding to genes (e.g. "md:lma_M00005") were included. This has been
corrected in the current version and thus may lead to different results
(2015/03/02).

```{r kegg_enrichment_modules, message=FALSE, warning=FALSE, results='hide', cache=CONFIG$use_cache, autodep=TRUE}
# Check each module for enrichment in KEGG terms and save result in a list
cl <- makeCluster(min(12, detectCores() - 1))
registerDoParallel(cl)

# Check each module for enrichment in GO terms and save result in a list
module_kegg_enrichment <- foreach(color=unique(module_colors), .packages=c('goseq')) %dopar% {
    # Measure KEGG enrichment for module
    enriched <- tryCatch({
        in_module_geneids <- gene_ids[module_colors == color]
        enriched <- test_gene_enrichment(in_module_geneids, gene_ids,
                                        gene_kegg_mapping, gene_lengths)
        # 2015/11/13
        # T. cruzi KEGG problem:
        #
        # Columns incorrectly names? (category saved as "description")
        # 
        #> kegg_pathways$description
        # [1] "path:tcr00900" "path:tcr01100" "path:tcr01110" "path:tcr01130" "path:tcr03010" "path:tcr00230" "path:tcr03050" "path:tcr04141" "path:tcr01040" "path:tcr01212"
        #[11] "path:tcr04146" "path:tcr00100" "path:tcr03015" "path:tcr05142" "path:tcr00330" "path:tcr03013" "path:tcr03040" "path:tcr00480" "path:tcr00220" "path:tcr00250"
        enriched <- unique(merge(enriched, kegg_pathways[,c('category','name')],
                                by='category'))
    }, error=function(e) {
        # goseq fails in some cases; have not been able to track down cause yet
        warning(sprintf("KEGG enrichment failed for module %s", color))
        return(get_enrichment_placeholder())
    })
    enriched
}

names(module_kegg_enrichment) <- unique(module_colors)

# remove any null entries from the results
module_kegg_enrichment <- module_kegg_enrichment[!sapply(module_kegg_enrichment, is.null)]

# unregister cpus
stopCluster(cl)
```

```{r print_kegg_enrichment_modules, results='asis'}
if (CONFIG$include_tables) {
    print_enrichment_results(module_kegg_enrichment, module_sizes, 'KEGG pathway',
                                kegg_pathways %>% select(-description),
                                'output/modules', 'kegg')
}

enriched_colors_kegg <- get_enriched_modules(module_kegg_enrichment)

# save results
COEXPRESSION_NETWORK_RESULT$enriched_colors_kegg <- enriched_colors_kegg
COEXPRESSION_NETWORK_RESULT$module_kegg_enrichment <- module_kegg_enrichment
COEXPRESSION_NETWORK_RESULT$kegg_pathways <- kegg_pathways
```

```{r enrichment_status}
go_enrichment_status   <- as.numeric(module_colors %in% enriched_colors_go)
kegg_enrichment_status <- as.numeric(module_colors %in% enriched_colors_kegg)
```

```{r module_dendrogram_annotated, results='asis', echo=CONFIG$include_plots, eval=CONFIG$include_plots}

cat('\n### Dendrogram with annotated modules\n')

# testing 2014/11/07
unassigned_modules <- as.numeric(module_colors == 'grey')

plotDendroAndColors(gene_tree,
                    cbind(module_colors, go_enrichment_status,
                          kegg_enrichment_status, unassigned_modules),
                    groupLabels=c(sprintf("Modules\n(n=%s)", num_modules),
                                  #sprintf("Red = upregulated at %s", CONFIG$de_cond2),
                                  "GO enrichment", "KEGG enrichment", "Unassigned"),
                    cex.colorLabels=cex_color_labels, cex.main=cex_main,
                    cex.axis=cex_axis, cex.lab=cex_lab,
                    dendroLabels=FALSE,
                    marAll=c(4,8,6,4),
                    guideHang=0.05)
```
