### Possible Sources for Comparison

- *L. infantum* exoproteome (Braga et al; 2014)
  - Manuscript includes a table of secreted proteins with TriTrypDB IDs.

Additional annotations for L. major
-----------------------------------

### Gene Conservation (dn/ds)

Mukherjee *et al* (2015) explored the properties of *Leishmania*-specific
"orphan genes". As part of their analyses, the average rate of non-synonymous
to synonymous substitutions ([dn/ds](http://en.wikipedia.org/wiki/Ka/Ks_ratio))
was calculated for each gene.

Below, we will use this information to look for possible links between
evolutionary conservation of protein-coding genes, and their topological
properties in the co-expression network.

**Reference:**

- S. Mukherjee, A. Panda, and T. C. Ghosh, “Elucidating evolutionary features
and functional implications of orphan genes in Leishmania major,” Infect.
Genet. Evol., vol. 32, pp. 330–337, 2015.

```{r lmajor_gene_conservation}
# Load supplementary dataset from Mukherjee et al (2015)
gene_conservation <- read_csv(file.path(input_dir, 'mukherjee_et_al_2015.csv'))

# Add centrality measures
#gene_conservation <- merge(merge(gene_conservation %>% select(gene_id, average_dnds),
#                          hub_scores, by='gene_id'),
#                          closeness_centrality, by='gene_id')
gene_conservation <- merge(gene_conservation %>% select(gene_id, average_dnds),
                          hub_scores, by='gene_id')

# hub genes
ggplot(gene_conservation, aes(average_dnds, hub_score)) + 
    geom_point(size=0.3) +
    geom_smooth()

# degree
gene_conservation <- merge(gene_conservation,
                          data.frame(gene_id=gene_ids,
                                     degree=rowSums(adjacency_matrix)),
                          by='gene_id')
ggplot(gene_conservation, aes(average_dnds, degree)) + 
    geom_point(size=0.3) +
    geom_smooth()

print(cor(gene_conservation$degree, gene_conservation$average_dnds))
print(cor(gene_conservation$hub_score, gene_conservation$average_dnds))
# closeness centrality
#ggplot(gene_conservation, aes(average_dnds, closeness)) + geom_point()
```

### Metabolic roles (Opperdoes et al. 2007)

(Not currently being used, but keeping for future reference...)

```{r lmajor_opperdoes2007}
# Add column indicicating plausible metabolic role for various genes; based
# on a supplementary table in Opperdoes et al (2007)
metabolic_role <- read.csv(file.path(input_dir, 'opperdoes2007_clean.csv'),
                          sep='\t', header=TRUE, na.strings="")

# keep only the gene_id and functionX columns
metabolic_role <- metabolic_role[,c(1,4,5,6)]

# discard genes that have already been filtered out
metabolic_role <- metabolic_role[metabolic_role$gene_id %in% result$gene_id,]

# bind to annotation data frame
result <- plyr::join(result, metabolic_role, by='gene_id')
```

### Other possible annotation sources

- http://onlinelibrary.wiley.com/doi/10.1038/msb.2009.13/full (T. cruzi; bionformatic
  & experimental -- no gene lists though)

