```{r create_output_directories, include=FALSE, message=FALSE, warning=FALSE}
# create output directory
timestring <- format(Sys.time(), "%Y%m%d")

# output directory    
output_dir <- file.path(CONFIG$output_dir, 
                        sprintf("%s_%s", CONFIG$analysis_name, timestring))
outdir <- file.path(output_dir, 'output')

# Directory to store module enrichment results to
module_output_dir <- file.path(outdir, 'modules')

if (!dir.exists(module_output_dir)) {
	dir.create(module_output_dir, recursive=TRUE)
}
```

General dendrogram features
---------------------------

Here we will explore the relationship between co-expression status and general
features such as expression variability and net change in expression across
time.

QUESTION: Can the variance between replicates for a given gene tell us anything
useful that with respect to network construction? For example, perhaps genes
with less variance within replicates are more tightly regulated. On the flip
side, genes with a larger variance may just be noise and not as interesting in
the context of the network.

```{r combine_replicates}
# expression averaged across replicates
combined_counts <- combine_replicates(exprs(network_counts$final), condition)
combined_counts_table <- cbind(gene_id=rownames(network_counts$final),
                              combined_counts)
```

```{r dendrogram_net_expression_change}
# Net change in expression over time
net_expr <- combined_counts[,CONFIG$main_comparison[2]] - combined_counts[,CONFIG$main_comparison[1]]

# general expression trends
sum(net_expr < 0)
sum(net_expr > 0)

# Threshold to avoid having extremes dominate color scale
net_expr_clip_lower <- as.numeric(quantile(net_expr, 0.01))
net_expr_clip_upper <- as.numeric(quantile(net_expr, 0.99))

net_expr <- pmax(net_expr_clip_lower, pmin(net_expr_clip_upper, net_expr))

# Shift to all positive range
midpoint <- abs(min(net_expr))
midpoint_percentile <-  round(sum(net_expr <= 0) / length(net_expr) * 100)

net_expr <- net_expr + midpoint

#pal <- colorRampPalette(brewer.pal(n=11, name='RdBu'))(100)
pal <- c(colorRampPalette(brewer.pal(n=9, name='Blues'))(midpoint_percentile),
        colorRampPalette(brewer.pal(n=9, name='Reds'))(100 - midpoint_percentile))
net_expr_colors <- pal[pmax(1, round((net_expr / max(net_expr)) * 100))]

# Gene std deviations
gene_std_devs <- log1p(sqrt(apply(exprs(network_counts$final), 1, var)))

gene_std_devs_clip_lower <- as.numeric(quantile(gene_std_devs, 0.01))
gene_std_devs_clip_upper <- as.numeric(quantile(gene_std_devs, 0.99))

gene_std_devs <- pmax(gene_std_devs_clip_lower, 
                     pmin(gene_std_devs_clip_upper, gene_std_devs))

pal <- colorRampPalette(brewer.pal(n=9, name='Greens'))(100)
std_dev_colors <- pal[pmax(1, round((gene_std_devs / max(gene_std_devs)) * 100))]

# Include main DE contrast
de_contrast <- lifecycle_stages[[main_contrast]]

# If no DE genes for plot, add blank row
if (is.null(de_contrast)) {
    de_contrast <- rep(0, length(module_colors))
}

# Dendrogram colors and labels
dendrogram_colors <- cbind(module_colors, de_contrast, net_expr_colors, std_dev_colors)
dendrogram_labels <- c(dendrogram_module_labels,
                       sprintf("Red = upregulated\n at %s", "72 hours"),
                       'Net expression', 'Log standard deviation')

#sprintf("%s vs. %s\n Red = Up-regulated\n Black = Down-regulated",
if (CONFIG$include_plots) {
    plotDendroAndColors(gene_tree, dendrogram_colors,
                        groupLabels=dendrogram_labels, dendroLabels=FALSE,
                        cex.colorLabels=cex_color_labels, cex.main=cex_main,
                        cex.axis=cex_axis, cex.lab=cex_lab, marAll=c(4,8,6,4),
                        guideHang=0.05)
}

```

