```{r knitr_settings, include=FALSE, message=FALSE, warning=FALSE}
library('knitr')
opts_knit$set(progress=FALSE, verbose=TRUE)
opts_chunk$set(error=FALSE)
options(digits=4)
options(stringsAsFactors=FALSE)
options(knitr.duplicate.label='allow')

# If rmarkdown.pandoc.to not specified (for example, when kniting
# piece-by-piece in Vim-R), have it default to 'latex' output.
if (is.null(opts_knit$get("rmarkdown.pandoc.to"))) {
    opts_knit$set(rmarkdown.pandoc.to='latex')
}

rm(list=ls())    # Clean up any existing variables
```

Libraries
---------

Let's start by loading all of the libraries that will be used throughout this
analysis.

```{r random_seed}
# Make sure results are reproducible
set.seed(1)
```

```{r child='includes_common.Rmd'}
```

```{r echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
# clear any existing bibliography
cleanbib()
```


Parameters
----------

In order to simplify the process of rerunning the following analysis with
different parameters, I have tried to abstract out many of the useful
experimental design parameters, enabling them to all be adjusting in one
location.

The parameters are broken up into two parts: first are the general settings
which apply to all analyses, regardless of species. The second set of settings
are those which are specific to the species being analysed.

```{r child='../settings/shared.Rmd'}
```

```{r child='../settings/differential_expression.Rmd'}
```

