Batch Adjustment
----------------

### Variable correlations (before batch correction)

```{r pca_feature_cor_before, results='asis'}
# Sample condition and covariate correlations with PCs
if (CONFIG$include_tables) {
    xkable(head(pca_feature_correlations(exprs(counts$normed), 
                                    cbind(condition, covariates)),
                                    caption=paste('Variable correlations with ',
                                                'principle components prior ',
                                                'to batch adjustment'), 10))
}
```

```{r apply_batch_adjustment, results='asis'}
# Create a new variable to store normalized/batch-adjusted counts
counts$batch_adjusted <- counts$normed

if (batch_adjust == 'limma') {
    cat(knit_child('batch_adjustment_limma.Rmd'))
} else if (batch_adjust == 'combat') {
    cat(knit_child('batch_adjustment_combat.Rmd'))
}
```

### Variable correlations (after batch correction)

```{r pca_feature_cor_after, results='asis', error=FALSE}
if (CONFIG$include_tables) {
    xkable(head(pca_feature_correlations(exprs(counts$batch_adjusted), 
                                    cbind(condition, covariates)),
                                    caption=paste('Variable correlations with ',
                                                'principle components ',
                                                'following batch adjustment'), 10))
}
```

### Sample plots before/after batch adjustment

#### Sample heatmap (normalized euclidean distance, before/after batch adjustment)

```{r sample_heatmap_dist_batch_adj, out.width=combined_plot_width, fig.show='hold', eval=CONFIG$include_plots}
plot_sample_heatmap(exprs(counts$normed), condition, batch,
                    main='Sample heatmap (euclidean distance, before batch adjustment)')
plot_sample_heatmap(exprs(counts$batch_adjusted), condition, batch,
                    main='Sample heatmap (euclidean distance, after batch adjustment)')
```

#### Sample heatmap (normalized pearson correlation, before/after batch adjustment)

```{r sample_heatmap_correlation_batch_adj, out.width=combined_plot_width, fig.show='hold', eval=CONFIG$include_plots}
plot_sample_heatmap(exprs(counts$normed), condition, batch,
                    metric='pearson',
                    main='Sample heatmap (pearson correlation, before batch adjustment)')
plot_sample_heatmap(exprs(counts$batch_adjusted), condition, batch,
                    metric='pearson',
                    main='Sample heatmap (pearson correlation, after batch adjustment)')
```

#### Sample PCA (normalized counts, before/after batch adjustment)

```{r sample_pca_before_after_batch_adjustment, out.width=combined_plot_width, fig.show='hold', eval=CONFIG$include_plots}
layout(t(1:2))
plot_sample_pca(exprs(counts$normed), sample_ids, condition,
                batch, main='Sample PCA (before batch adjustment)')
plot_sample_pca(exprs(counts$batch_adjusted), sample_ids, condition,
                batch, main='Sample PCA (after batch adjustment)')
```

```{r include=CONFIG$debug, eval=CONFIG$debug}
sum(exprs(counts$batch_adjusted))
```

```{r}
# For DEA, our "final" version of the counts is just the batch-adjusted count
# table
counts$final <- counts$batch_adjusted
```

