Normalization (Network)
-----------------------

Because the optimal parameters used for processing count data for co-expression
network construction may different from that used for the differential 
expression analysis performed previously, we will begin by reprocessing the
data and storing alternate version of each level of processing, specifically
for use during network construction.

Depending on the focus of the co-expression network, we may also drop some of
the samples to only retain samples for the specific conditions we are
interested in. Instead of limiting ourselves to these samples right away, we
will wait until a later step in order to preserve as much information as
possible for variance estimation and batch adjustment.

```{r pre_normalization}
# Create network version of counts
network_counts <- list(raw=de_counts$raw,
                       normed=de_counts$raw,
                       log2cpm=de_counts$log2cpm)

# Co-expression network analysis parameters
counts             <- network_counts
batch_adjust       <- CONFIG$network_batch_adjust
quantile_normalize <- CONFIG$network_quantile_normalize
use_cpm            <- CONFIG$network_cpm
use_log2           <- CONFIG$network_log2
use_voom           <- CONFIG$network_voom
robust_ebayes      <- CONFIG$network_ebayes_robust
```

### Voom, log2CPM, and quantile normalization

```{r child='normalize_counts.Rmd'}
```

```{r child='sample_visualization.Rmd'}
```

```{r batch_adjustment}
if (CONFIG$network_batch_adjust != 'none') {
    cat(knit_child('batch_adjustment.Rmd'))
}
```

```{r post_normalization}
# Store transformed network counts
network_counts <- counts

# Update condition, batch, etc. to include only network-specific samples
condition <- factor(CONFIG$wgcna_samples[[CONFIG$condition]])
batch     <- factor(CONFIG$wgcna_samples[[CONFIG$batch]])

# Samples to use for network construction
sample_ids_network  <- as.character(CONFIG$wgcna_samples[[CONFIG$sample_id]])
network_sample_mask <- sampleNames(network_counts$raw) %in% sample_ids_network

covariates <- covariates[network_sample_mask,]

# drop samples not used in network construction
network_counts$batched_adjusted <- network_counts$batch_adjusted[,network_sample_mask]
network_counts$final            <- network_counts$final[,network_sample_mask]
```

```{r include=CONFIG$debug, eval=CONFIG$debug}
sum(exprs(network_counts$final))
```

```{r child='network_filtering.Rmd'}
```

```{r update_network_counts}
# Update gene annotations
gene_info <- gene_info[gene_info$gene_id %in% rownames(network_counts$final),]

# Update rows and columns for each version of de_s and network_counts to only
# include those genes that remain in network_counts$final
gene_ids <- rownames(network_counts$final)

for (name in names(network_counts)) {
    # Update de count tables (batch_adjusted may not exist)
    if (name %in% names(de_counts)) {
        de_counts[[name]] <- de_counts[[name]][rownames(de_counts[[name]]) %in% gene_ids,]
    }
    # Update network count tables
    if (ncol(network_counts[[name]]) == length(network_sample_mask)) {
        sample_ind <- network_sample_mask
    } else {
        sample_ind <- rep(TRUE, ncol(network_counts[[name]])) 
    }
    network_counts[[name]] <- network_counts[[name]][rownames(network_counts[[name]]) %in% gene_ids, sample_ind]
}
```

```{r network_sample_vis, echo=FALSE, results='asis'}
# If the samples used for DEA and network construction differ, show the
# sample heatmap and PCA plots once more for only the network-specific samples
samples_differ <- length(sample_ids) == length(sample_ids_network)

if (samples_differ) {
    cat("\n## Network sample visualization\n\n")
    counts <- network_counts
}
```

```{r child='sample_visualization.Rmd', echo=samples_differ, eval=samples_differ}
```

