Filter data
-----------

Next let's filter the data to remove unwanted or uninformative genes.
Additionally, multicopy genes can be filtered at this step. Later on in the
analysis we may do some additional filtering based on expression variance (for
genes) or outlier analysis (for samples).

Filtering targets:

1. low-count genes
2. multicopy genes (optional)
3. noncoding RNAs

Later on (post-normalization), there may also be an addition step of
variance-based filtering applied.

### Remove low-count genes

```{r remove_lowcount_genes}
# for relatively small datasets, we can look for genes with non-zero expression in at least
# as many samples are contained by the smallest condition group
reps_per_batch <- table(factor(condition))
min_replicates <- min(reps_per_batch)

# in some cases, very few replicates (possibly only a single one) are present;
# we will use a lower limit of 3 samples to ensure that at least a few of the samples
# pass the specific low count threshold
min_samples_passing_threshold <- max(min_replicates, 3)

# for larger datasets with many conditions (e.g. > 1,000 samples), this may be too conservative,
# so we will also set a lower limit equal to 5% times the total number of samples
min_samples_passing_threshold <- max(min_samples_passing_threshold, round(length(condition) / 20))

de_counts$raw <- filter_lowcount_genes(de_counts$raw, 
                                       threshold=CONFIG$low_count_threshold,
                                       min_samples=min_samples_passing_threshold, verbose=TRUE)
```

```{r include=CONFIG$debug, eval=CONFIG$debug}
sum(exprs(de_counts$raw))
```

### Filter noncoding RNAs

Next, we will remove ncRNAs such as rRNAs, tRNAs, and snoRNAs, which are not
regulated in the same manner as other coding genes. Since these are not
polyadenylated, they should not be present in the samples in large numbers.

```{r filter_noncoding_rnas}
# Filter by gene id or type
# Note that when filtering by type for T. cruzi, SRPs and SLRNAs are not included
ncrna_mask <- rep(FALSE, nrow(de_counts$raw))

if (!is.null(CONFIG$id_filter_string)) {
    ncrna_mask <- ncrna_mask | grepl(CONFIG$id_filter_string,
                                     rownames(de_counts$raw))
}

if (!is.null(CONFIG$type_filter_string)) {
    gene_types <- (gene_info %>% filter(gene_id %in% rownames(de_counts$raw)))$type
    ncrna_mask <- ncrna_mask | grepl(CONFIG$type_filter_string, gene_types)
}

num_noncoding_rnas <- sum(ncrna_mask)
print(sprintf("Removing %d non-coding RNAs", num_noncoding_rnas))
de_counts$raw <- de_counts$raw[!ncrna_mask,]
```

```{r trim_gene_info}
# Keep annotations for remaining genes only
gene_info <- gene_info[gene_info$gene_id %in% rownames(de_counts$raw),]
```

```{r highly_expressed_genes_after, results='asis', include=!CONFIG$filter_multicopy, eval=!CONFIG$filter_multicopy}
cat('\n### Highly expressed genes (after filtering)\n')

num_reads_total <- sum(as.numeric(exprs(de_counts$raw)))

gene_total_counts <- data.frame(sort(rowSums(exprs(de_counts$raw)), decreasing=TRUE))
colnames(gene_total_counts) <- c("total_reads")
gene_info_subset <- gene_info[match(rownames(gene_total_counts),
                                    gene_info$gene_id),]
gene_info_subset$num_reads <- gene_total_counts$total_reads
gene_info_subset$num_reads_pct <- (gene_total_counts$total_reads / num_reads_total) * 100

gene_info_subset <- gene_info_subset %>% 
    dplyr::select(-chromosome, -strand) %>%
    dplyr::rename(tx_len=transcript_length)

if (CONFIG$include_tables) {
    xkable(head(gene_info_subset, 10), 
        caption='Genes with the highest average expression after filtering.',
        str_max_width=20)
}
```

```{r log2cpm_counts}
# Create copies of raw count ExpressionSet
de_counts$log2cpm <- de_counts$raw

# Keep a copy of the Log2-CPM normalized counts for plotting purposes
exprs(de_counts$log2cpm) <- log2(counts_per_million(exprs(de_counts$raw)) + 0.5)

# clean-up
suppressMessages(gc())
```

```{r include=CONFIG$debug, eval=CONFIG$debug}
sum(exprs(de_counts$raw))
```

