### Network construction

#### Overview

Next we will construct a co-expression network using WGCNA. There are a large
number of parameters that go in to the construction.

#### Construct similarity matrix

TODO: Update description.

First, we construct a similarity matrix from the gene expression profiles.

One simple way to measure the similarity between gene expression profiles is to
take the pearson correlation of the each pair of profiles:

$$
s_{ij} = cor(x_i,x_j)
$$

where $x_i$ and $x_j$ are the expression profiles corresponding with genes $i$
and $j$.

```{r construct_similarity_matrix}
message("Constructing co-expression network")

# WGCNA count input
wgcna_input <- exprs(network_counts$final)

# Convert to float (necessary for cor)
wgcna_input <- wgcna_input * 1.0

# Pearson correlation / bi-midweight correlation
if (CONFIG$similarity_measure %in% c('cor', 'bicor')) {
    sim_measure <- get(CONFIG$similarity_measure)

    if (CONFIG$similarity_measure == 'cor') {
        similarity_matrix <- sim_measure(t(wgcna_input))
    } else if (CONFIG$similarity_measure == 'bicor') {
        # maxPOutliers discussion
        # https://support.bioconductor.org/p/65124/
        similarity_matrix <- sim_measure(t(wgcna_input), maxPOutliers=0.1)
    }
} else if (CONFIG$similarity_measure == 'spearman') {
    # Spearman correlation
    similarity_matrix <- cor(t(wgcna_input), method='spearman')

    # 2015/12/19 Spearman sometimes returns values outside of [-1, 1]
    # see: https://bugs.r-project.org/bugzilla3/show_bug.cgi?id=16638
    similarity_matrix[similarity_matrix < -1] <- -1
    similarity_matrix[similarity_matrix > 1] <- 1
} else if (CONFIG$similarity_measure == 'cor-dist') {
    # Combination of Pearson correlation and Euclidean distance
    cor_matrix  <- cor(t(wgcna_input))
    dist_matrix <- as.matrix(dist(wgcna_input, diag=TRUE, upper=TRUE))
    dist_matrix <- log1p(dist_matrix)
    dist_matrix <- 1 - (dist_matrix / max(dist_matrix))

    similarity_matrix <- sign(cor_matrix) * ((abs(cor_matrix) + dist_matrix)/ 2)

    # clean-up
    rm(cor_matrix)
    rm(dist_matrix)
    gc()
} else {
    # Euclidean distance

    # Old way of converting euclidean distance to a similarity measure:
    # inverse of max-scaled log-euclidean-distance
    similarity_matrix <- as.matrix(dist(wgcna_input, diag=TRUE, upper=TRUE))
    similarity_matrix <- log1p(similarity_matrix)
    similarity_matrix <- 1 - (similarity_matrix / max(similarity_matrix))

    # Gaussian kernel distance
    # Produces a matrix in the range [0,1]
    # http://www.inside-r.org/packages/cran/KRLS/docs/gausskernel
    #sigma <- median(rowSums(wgcna_input**2))
    #similarity_matrix <- exp(-1 * as.matrix(dist(wgcna_input)^2)/sigma)

    if (CONFIG$similarity_measure == 'dist') {
        # For convenience, modify to range from [0,1] to [-1,1]; this will allow 
        # downstream steps that expect output from cor() to behave the same
        similarity_matrix <- (2 * similarity_matrix) - 1
    } else if (CONFIG$similarity_matrix == 'signed-dist') {
        similarity_matrix <- sign(cor(t(wgcna_input))) * similarity_matrix
    }
}

# Preserve sign of correlations
# (not currently used...)
#neg_correlations <- similarity_matrix < 0

# For unsigned networks, take the absolute value of the correlations
if (CONFIG$network_type == 'unsigned') {
    similarity_matrix <- abs(similarity_matrix)
}
```

```{r include=CONFIG$debug, eval=CONFIG$debug}
sum(similarity_matrix)
```

```{r similarity_matrix_density, message=FALSE, cache=CONFIG$use_cache, cache.lazy=FALSE, autodep=TRUE, fig.width=CONFIG$fig_height/CONFIG$dpi}
if (CONFIG$include_plots) {
    plot(density(similarity_matrix))
}
```

```{r similarity_matrix_heatmap, message=FALSE, cache=CONFIG$use_cache, cache.lazy=FALSE, autodep=TRUE, fig.width=CONFIG$fig_height/CONFIG$dpi, dev='png'}
# Similarity matrix plot, downsampled to speed things up
heatmap_indices <- sample(nrow(similarity_matrix), min(1600, nrow(similarity_matrix)))

if (CONFIG$include_plots) {
    heatmap.2(t(similarity_matrix[heatmap_indices, heatmap_indices]),
                hclustfun=flashClust, col=redgreen(75), margin=c(7,7),
                labRow=NA, labCol=NA, 
                key.xlab=sprintf('Similarity\n(%s)', CONFIG$similarity_measure),
                trace='none',
                xlab='Gene', ylab='Gene',
                #lwid=c(1, 4), lhei=c(1, 4),
                cexRow=2, cexCol=2, keysize=1,
                main='Similarity matrix',
                density.info='none', revC=TRUE)
}

# Set the diagonal to zero to remove uninformative correlations
diag(similarity_matrix) <- 0

```

#### Construct adjacency matrix

Convert similarity matrix to an adjacency matrix.

In the case of a signed network, in order to preserve the sign of the
correlations, WGCNA's `adjacency()` function also shifts and scales the data
from the range [-1,1] to [0,1]:

$$
a_{ij} = \left(\frac{1}{2} (1 + s_{ij})\right)^\gamma
$$

This enables us to use the topological overlap matrix functions downstream
which expect signed correlations in this format (providing these functions with
data in the range [-1,1] will not behave well, even when the functions are told
to expect signed input.)

Some additional reading regarding the use of power transformation:

1. [http://labs.genetics.ucla.edu/horvath/CoexpressionNetwork/Rpackages/WGCNA/faq.html](WGCNA FAQ) (E.g. item #5)
2. [https://support.bioconductor.org/p/66101/#66195](Bioconductor Support - Question: Theoretical WGCNA Question)

```{r construct_adjacency_matrix, cache=CONFIG$use_cache, cache.lazy=FALSE, autodep=TRUE}
# Construct adjacency matrix
adjacency_matrix <- adjacency.fromSimilarity(similarity_matrix,
                                            power=CONFIG$adjmatrix_power,
                                            type=CONFIG$network_type)

# Delete similarity matrix to free up memory
rm(similarity_matrix)
gc()

# Convert to matrix
adjacency_matrix <- matrix(adjacency_matrix, nrow=nrow(adjacency_matrix))
rownames(adjacency_matrix) <- gene_ids
colnames(adjacency_matrix) <- gene_ids
```

```{r adjacency_matrix, message=FALSE, cache=CONFIG$use_cache, cache.lazy=FALSE, autodep=TRUE, fig.width=CONFIG$fig_height/CONFIG$dpi, dev='png'}
# scale colormap to improve contrast
#green <- redgreen(10)[8]
#cmap <- c(colorpanel(30, '#FF0000', '#000000', green),
#            colorpanel(70, green, '#00FF00'))
diag(adjacency_matrix) <- 1

if (CONFIG$include_plots) {
    hmap <- heatmap.2(adjacency_matrix[heatmap_indices, heatmap_indices],
            hclustfun=flashClust, col=redgreen(75),
            labRow=NA, labCol=NA, trace='none', revC=TRUE,
            xlab='Gene', ylab='Gene', density.info='none',
            main='Adjacency matrix')
    # save column and row order for future plotting
    heatmap_row_idx <- hmap$rowInd
    heatmap_col_idx <- hmap$colInd
}
diag(adjacency_matrix) <- 0
```

```{r include=CONFIG$debug, eval=CONFIG$debug}
sum(adjacency_matrix)
```

Done! We now have a co-expression network to analyze. The next step is to
cluster the genes in order to detect expression modules within the network.

